### 2024-10-28 paper/3-motivation.tex
Explain with rewrite rule

% control flow structure in SpecTec prose
However, the \spectecp{} assumes different a model to explain the control flow
of WebAssembly.
Rather than using the notion of a pc, \spectecp{} assumes that WebAssembly
instructions are just given one by one.
This is because the \spectecp{} is generated automatically from the
\red{SpecTec dsl}.
\red{SpecTec dsl} uses rewrite rule to describe WebAssembly semantics, and it
can be seen as consuming a WebAssembly instruction and values in the stack,
changing the stack and the store, and inputing new WebAssembly instructions.

@@

Consider example 1 above.

\texttt{loop} and \texttt{unreachable} is the input instruction for the model.
\cref{fig:spectec-loop} is the \spectecp{} of the \texttt{loop} instruction.
Rather than storing the point to jump as a continuation in the label, it stores
the \texttt{loop} instruction itself and \textit{enters} the block.
Here, \textit{enter} means that it pushes the label in the stack and inputs the
instructions as inputs to execute.
As a result, after the \texttt{loop} is executed, \texttt{i32.const},
\texttt{br}, \texttt{unreachable}, and \texttt{unreachable} are remained as
inputs to the model.
When \texttt{i32.const} is excuted, it pushes 42 as \officialp{} does.
However, \texttt{br} behaves a bit differently.
\cref{fig:spectec-br} is the \spectecp{} of the \texttt{br} instruction.
It pops the label from the stack, removes the input instructions until the end
of the block, considers the loop instruction in the label as a new input
instruction, and executes it again.
Consequently, remaining inputs are \texttt{loop}, and \texttt{unreachable}.
Therefore, it explains the behavior in different point of view, but actually
describes the same thing.

% exiting label in SpecTec prose
However, this viewpoint is challenged by the exiting label.
As the block structure doesn't remain intact, it is hard to describe
\textbf{the end of the block}.
Furthermore, the exiting label is not performed by a specific WebAssembly
instruction, which makes it hard to model the behavior.
To handle this problem, an administrative instruction \texttt{end} is
introduced in \spectecp.
When entering a block, assume that there is an \texttt{end}

>>>>

% control flow structure in SpecTec prose
However, the \spectecp{} assumes a different model to explain the control flow
of WebAssembly.
Rather than using the notion of a pc, \spectecp{} assumes that WebAssembly
instructions are just given one by one.
This is because the \spectecp{} is generated automatically from the
\red{SpecTec DSL}.
\red{SpecTec DSL} uses rewrite rule to describe WebAssembly semantics, and it
can be seen as consuming a WebAssembly instruction, pushing or poping a control
structures, and inputing new WebAssembly instructions.
It can be expressed in the following form:
$c^* \vdash i_0, i_1, ..., i_m \leadsto c'^* \vdash i_1, ..., i_m$

% challenges in SpecTec prose
However, this viewpoint is challenged by the exiting label.
As the block structure doesn't remain intact in the model, it is hard to
describe \textbf{the end of the block}.
Furthermore, the exiting label is not performed by a specific WebAssembly
instruction, which makes it hard to model the behavior.
To handle this problem, an administrative instruction \texttt{end} is
introduced in \spectecp.
When entering a block, an \texttt{end} is added to express the end of the block.

@@

\begin{align}
  &c^*
  \vdash
  loop(i32, ~ [ ~ i32.const ~ 42, ~ br ~ 0, ~ unreachable ~ ]), ~ unreachable
\leadsto \\
  &label(loop(i32, ~ [ ~ i32.const ~ 42; ~ br ~ 0; ~ unreachable ~ ])), c^*
  \vdash
  i32.const ~ 42, ~ br ~ 0, ~ unreachable, ~ end, ~ unreachable
\leadsto \\
  &label(loop(i32, ~ [ ~ i32.const ~ 42; ~ br ~ 0; ~ unreachable ~ ])), c^*
  \vdash
  ~ br ~ 0, ~ unreachable, ~ end, ~ unreachable
\leadsto \\
  &c^*
  \vdash
  loop(i32, ~ [ ~ i32.const ~ 42; ~ br ~ 0; ~ unreachable ~ ]), unreachable \\
\end{align}

Consider example 1 above.
If we assume some control structure $c^*$ is given, we can express the code
like (3.1).
\cref{fig:spectec-loop} is the \spectecp{} of the \texttt{loop} instruction.
Rather than storing the point to jump as a continuation in the label, it stores
the \texttt{loop} instruction itself and \textit{enters} the block.
Here, \textit{enter} means that it pushes the label in the stack and executes
the instructions in the block, which means that \texttt{i32.const},
\texttt{br}, \texttt{unreachable}, \texttt{end} and \texttt{unreachable} become
the new inputs (3.2).
When \texttt{i32.const} is excuted, it pushes 42 as \officialp{} does (3.3).
The value is omitted for brevity.
However, \texttt{br} behaves a bit differently.
\cref{fig:spectec-br} is the \spectecp{} of the \texttt{br} instruction.
It pops the label from the stack, removes the input instructions until the end
of the block, considers the loop instruction in the label as a new input
instruction.
Consequently, remaining inputs are \texttt{loop}, and \texttt{unreachable}
again.
Therefore, it explains the same behavior in different point of view.
